---
title: "Homework 1"
subtitle: "Statistical Rethinking"
date: "`r strftime(Sys.Date(), format = '%Y-%m-%d')`"
author: Patrick Anker
header-includes:
output: 
  bookdown::pdf_document2:
    includes:
      in_header: "../preamble.tex"
    toc: false
    extra_dependencies: "float"
    number_sections: false
---

```{r setup}
library(ggplot2)
set.seed(0xABBA)

line_color <- "#d12e66"
line_color2 <- "#027ea5"

hist_fill <- "#dadada"
hist_color <- "grey"
```

# Question 1

_Suppose the globe tossing data had turned out to be 4 water and 11 land. Construct the posterior distribution._

I'll continue to use the flat prior for $p$.

\begin{align*}
W\sim \mathrm{Binom}(N = 15, p) \\
p\sim \mathrm{Uniform}(0, 1)
\end{align*}

The likelihood is given as

$$
P(W = 4, L = 11 | p) = \binom{15}{4}p^4(1-p)^{11}
$$

Via grid approximation,

```{r}
n <- 100
p_grid <- seq(from = 0, to = 1, length.out = n)
prob_p <- rep(1, n)

post <- dbinom(4, size = 15, prob = p_grid) * prob_p
post <- post / sum(post)

dat <- tidytable::tidytable(p = p_grid, posterior = post)

ggplot(dat, aes(x = p, y = posterior)) +
  geom_point() +
  theme_bw()
```

# Question 2

_Using the posterior distribution from **1**, compute the posterior predictive for the next 5 tosses of the same globe._

Creating samples from the posterior:

```{r}
n_samps <- 1e4
samples <- sample(p_grid, size = n_samps, replace = TRUE, prob = post)
```

Now to examine the next few tosses:

```{r}
pops_w <- tidytable::tidytable(
  w = rbinom(n_samps, size = 5, prob = samples),
  sample = seq_len(n_samps)
)

ggplot(pops_w, aes(x = w)) +
  geom_bar(fill = hist_fill, color = hist_color) +
  theme_bw()
```

# Question 3

_Use the posterior predictive distribution from **2** to calculate the probability of 3 or more water samples in the next 5 tosses._

```{r}
pops_w |>
  tidytable::summarise(
    three_or_more = max(w) >= 3,
    .by = sample
  ) |>
  tidytable::pull(three_or_more) |>
  mean()
```

# Question 4

_This problem is an optional challenge for people who are taking the course for a second or third time. Suppose you observe $W = 5$ water points, but you forgot to write down how many times the globe was tossed, so you don’t know the number of land points $L$. Assume that $p = 0.7$ and compute the posterior distribution of the number of tosses $N$. Hint: Use the binomial distribution._

Normally, I would use a NegBinom distribution here, but I will go with the Binom distribution and vary $N$. We know $N\geq 5$. I will use a Poisson distribution to model the prior of $N$ with 0 probability density less than 5.

```{r, out.width="65%", fig.align="center", fig.cap="Comparison of prior and posterior distributions of $N$"}
n_grid <- 0:30

prior_n <- dpois(n_grid, lambda = 10)
prior_n[n_grid < 5] <- 0
prior_n <- prior_n / sum(prior_n) # Normalization for vec to be probabilities

post_n <- dbinom(5, size = n_grid, prob = 0.7) * prior_n
post_n <- post_n / sum(post_n)

dat <- tidytable::tidytable(
  N = c(n_grid, n_grid),
  density = c(prior_n, post_n),
  distribution = factor(c(
    rep("Prior", length(n_grid)),
    rep("Posterior", length(n_grid))
  )),
)

ggplot(dat, aes(x = N, y = density, color = distribution)) +
  geom_point() +
  geom_line(alpha = 0.5) +
  theme_bw()
```

